---
title: 'Final Project: Housing Data'
author: "Jacob Minkin"
date: "May 19, 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
## Intro

We start this project by loading the data from GitHub.
```{r setup}
rm(list = ls())
pacman::p_load(tidyverse,data.table, R.utils, stringr,missForest, YARF, mlr )
options(java.parameters = "-Xmx8000m")
housing_data = fread("https://raw.githubusercontent.com/kapelner/QC_MATH_342W_Spring_2021/master/writing_assignments/housing_data_2016_2017.csv")

Northeast_Queens_zip = c(11361, 11362, 11363, 11364)
North_Queens_zip = c(11354, 11355, 11356, 11357, 11358, 11359, 11360)
Central_Queens_zip = c(11365, 11366, 11367)
Jamaica_zip = c(11412, 11423, 11432, 11433, 11434, 11435, 11436)
Northwest_Queens_zip = c(11101, 11102, 11103, 11104, 11105, 11106)
West_Central_Queens_zip = c(11374, 11375, 11379, 11385)
Southeast_Queens_zip = c(11004, 11005, 11411, 11413, 11422, 11426, 11427, 11428, 11429)
Southwest_Queens_zip = c(11414, 11415, 11416, 11417, 11418, 11419, 11420, 11421)
West_Queens_zip = c(11368, 11369, 11370, 11372, 11373, 11377, 11378)
```

## Remove Garbage

 First we are only predicting on data that includes the sale pricel. We remove all data that does not include the sale price.

 The important features start with 'approx_year_built'. Remove 'url', 'common_charges', 'model_type', 'date_of_sale', and 'listing_price_to_nearest_1000'. Also remove 'sq_footage' 'parking_charge' and 'total_taxes' because too much is missing.

```{r remove garbage}
no_garabge_data = housing_data %>% 
 # filter(!is.na(sale_price)) %>% 
  select(approx_year_built:last_col(), -url, -common_charges, -model_type, -date_of_sale, -listing_price_to_nearest_1000, -parking_charges)

table(no_garabge_data$coop_condo, exclude = NaN)
```

## Manipulating Data

Here is a list of manipulations
1. Pets allowed is a combo of cats and dogs allowed. if one pet is allowed then pets allowed is true.
2. fuel type is cleaned by combining others. Some were capitalized.
3. sq footage was factorized into small medium large super large and unknown.
4. Garage exists takes all the na's and makes them no. All others are yes.
5. Missing half bathrooms are set to zero.
6. In the dining_room_type, we assume 'dining area' is considered 'formal'. Set 'NA' as "unknown".
7. In the 'kitchen_type' feature, there was a observation from 'approx_year_built'. Added that data back and cleaned the rest of 'kitchen_type'
8. Converted 'maintenance_cost' and 'parking_charges' into numerics by removing the dollar sign and the comma.
7. Get zip_code from 'full_address_or_zip_code' and sort it into regions.
8. Made 'decade_built' from approx_year_built
9. Finally put 'sale_price' in front and remove 'zip_code', 'full_address_or_zip_code' and 'approx_year_built'.

```{r manipulations}
manipulated_data = no_garabge_data %>% 
  
  mutate(cats_allowed = if_else(cats_allowed == "y", "yes", cats_allowed)) %>% 
  mutate(dogs_allowed = if_else(dogs_allowed == "yes89", "yes", dogs_allowed)) %>%
  mutate(pets_allowed = if_else(dogs_allowed =="no" & cats_allowed == "no", "no", "yes")) %>% 
  
  mutate(fuel_type = if_else(fuel_type == "Other", "other", fuel_type)) %>% 
  
  mutate(sq_footage = as.factor(case_when(sq_footage %in% 100:600 ~ "small", sq_footage %in% 601:998 ~ "medium", sq_footage %in% 1000:1800 ~ "large", sq_footage %in% 1950:6215 ~ "super large", is.na(sq_footage) ~ "Unknown"))) %>% 
  
  mutate(garage_exists = if_else(is.na(garage_exists), "No", "Yes")) %>% 
  
  mutate(num_half_bathrooms = replace_na(num_half_bathrooms, 0)) %>% 
  
  mutate(dining_room_type = if_else(dining_room_type == "dining area", "formal", dining_room_type)) %>% 
  mutate(dining_room_type = ifelse(is.na(dining_room_type), "Unknown", dining_room_type)) %>%  
  
  mutate(approx_year_built = if_else(kitchen_type == "1955", as.integer(1955), approx_year_built, missing = approx_year_built)) %>%  
  mutate(kitchen_type = as.factor( case_when(kitchen_type == "combo" ~ "Combo", kitchen_type == "Combo" ~ "Combo", kitchen_type == "eat in" ~ "Eat In", ... = kitchen_type == "Eat in" ~ "Eat In", kitchen_type == "Eat In" ~ "Eat In", kitchen_type == "eatin" ~ "Eat In", str_sub(kitchen_type, start= 1, end = 3)=="eff"~ "efficiency" ))) %>% 

  mutate(maintenance_cost =  as.numeric(gsub("[\\$,\\,]", "", maintenance_cost))) %>% 
  mutate(sale_price =  as.numeric(gsub("[\\$,\\,]", "", sale_price))) %>% 

  mutate(zip_code = str_sub(full_address_or_zip_code, start= -5), zip_code = if_else(zip_code == "Share", as.integer(11354), as.integer(zip_code))) %>% 
  mutate(region = as.factor( case_when(zip_code %in% Northeast_Queens_zip  ~ "Northeast Queens", zip_code %in% North_Queens_zip  ~ "North Queens", zip_code %in% Central_Queens_zip  ~ "Central Queens", zip_code %in% Jamaica_zip  ~ "Jamaica", zip_code %in% Northwest_Queens_zip  ~ "Northwest Queens", zip_code %in% West_Central_Queens_zip  ~ "West Central Queens", zip_code %in% Southeast_Queens_zip  ~ "Southeast Queens", zip_code %in% Southwest_Queens_zip  ~ "Southwest Queens", zip_code %in% West_Queens_zip  ~ "West Queens"))) %>%  
  
  mutate(decade_built = as.factor(case_when(approx_year_built %in% 1915:1939 ~ "1915 - 1939", approx_year_built %in% 1940:1949 ~ "1940's", approx_year_built %in% 1950:1959 ~ "1950's", approx_year_built %in% 1960:1969 ~ "1960's", approx_year_built %in% 1970:1979 ~ "1970's", approx_year_built %in% 1980:1989 ~ "1980's", approx_year_built %in% 1990:1999 ~ "1990's", approx_year_built %in% 2000:2009 ~ "2000's", approx_year_built %in% 2010:2017 ~ "2010's"))) %>% 
  
  
  
  select(sale_price, everything(), -full_address_or_zip_code, -zip_code, -approx_year_built)
  
  
 # table(manipulated_data$dining_room_type, exclude = NaN)
```

##co-op Condo cost
There is a big difference between co-op and condo. we will seperate them into 2 data frames and find the total cost. we will then combine the cases.
```{r}
condo_cases = manipulated_data %>% 
  filter(coop_condo == "condo") %>% 
  mutate(maintenance_cost = if_else(is.na(maintenance_cost), 0, maintenance_cost)) %>% 
  mutate(pct_tax_deductibl = if_else(is.na(pct_tax_deductibl),0, pct_tax_deductibl/100)) %>% 
  mutate(total_taxes = parse_number(total_taxes)) %>% 
  mutate(total_cost = maintenance_cost * (1-pct_tax_deductibl) + total_taxes)
  
  
coop_cases = manipulated_data %>% 
  filter(coop_condo == "co-op") %>% 
  mutate(pct_tax_deductibl = if_else(is.na(pct_tax_deductibl), mean(pct_tax_deductibl, na.rm = TRUE)/100, pct_tax_deductibl/100)) %>% 
  mutate(total_taxes = if_else(is.na(total_taxes), 0, parse_number(total_taxes))) %>% 
  mutate(total_cost = maintenance_cost * (1-pct_tax_deductibl)+ total_taxes)

data_with_cost = bind_rows(condo_cases, coop_cases) %>% 
  select(-maintenance_cost, - pct_tax_deductibl, -total_taxes, -cats_allowed, -dogs_allowed)
  
#table((data_with_cost$total_cost), exclude = NaN)
```


##Train and Test split
First we need to filter out the the missing sale prices. Then we will find the indicies of the split for train and test. 
```{r}
filtered_data = data_with_cost %>% 
  filter(!is.na(sale_price)) %>% 
  mutate_if(sapply(data_with_cost, is.character), as.factor)

missing_sale_price = data_with_cost %>% 
  filter(is.na(sale_price)) %>% 
  mutate_if(sapply(data_with_cost, is.character), as.factor)

data_summary = lapply(filtered_data, summary)
print (data_summary)
K = 4
test_indices = sample(1 : nrow(filtered_data), round(nrow(filtered_data) / K))
train_indices = setdiff(1 : nrow(filtered_data), test_indices)
test_data_miss = filtered_data[test_indices, ]
train_data_miss = filtered_data[train_indices, ]

miss_data_train_combined = bind_rows(train_data_miss, missing_sale_price) %>% 
  mutate(sale_price_dummy = if_else(is.na(sale_price), 0, 1))
miss_data_train_combined = miss_data_train_combined %>% 
  mutate_if(sapply(miss_data_train_combined, is.character), as.factor)

ximpMF = missForest(miss_data_train_combined)
train_data = ximpMF$ximp %>% 
  filter(sale_price_dummy == 1) %>% 
  select(-sale_price_dummy)

y_train = train_data$sale_price
X_train = train_data[ ,-1]
```

##Modeling
We will build 3 models. 1. Regression Tree 2. Linear Model 3. Random Forest

## Regression Tree
```{r}
tree_model = YARF(data.frame(x = X_train), y_train, num_trees = 1)
tree_model

```

## Linear Model
```{r}
linear_mod = lm(sale_price ~ ., train_data)
sd(y_train - linear_mod$fitted.values)
summary(linear_mod)
```
## Random  Forest
```{r}
mtry_mlr= tuneRF(X_train,
                y_train,
                stepFactor=0.5,
                plot=TRUE,
                ntreeTry=300,
                trace=TRUE,
                improve = 0.05)
print(mtry_mlr)

```

```{r}
mod_rf = YARF(X_train, y_train, mtry = 10)
mod_rf
```

##Performance of Random Forest

```{r}
y_test = test_data_miss$sale_price
x_test_miss = test_data_miss %>% 
  mutate(sale_price = -1)
miss_data_test_combined = bind_rows(train_data_miss, missing_sale_price, x_test_miss) %>% 
  mutate(sale_price_dummy = if_else(sale_price == -1, 1, 0)) %>% 
  mutate(sale_price_dummy = if_else(is.na(sale_price_dummy), 0, sale_price_dummy)) %>% 
  mutate(sale_price = na_if(sale_price, -1))
miss_data_test_combined = miss_data_test_combined %>% 
  mutate_if(sapply(miss_data_test_combined, is.character), as.factor)

test_imputed = missForest(miss_data_test_combined)

X_test = test_imputed$ximp %>% 
  filter(sale_price_dummy == 1) %>% 
  select(-sale_price, -sale_price_dummy)

y_hat = predict(mod_rf, X_test)

y_bar = mean(y_test)
SSR = sum((y_hat-y_bar)^2)
SST = sum((y_test-y_bar)^2)
rsq = (SSR/SST)
rsq
sd(y_test - y_hat)
```

